Standalone Tactics using OpenTheory

### 2012-02-07 paper.tex
bit more paper hacking

An article file represents a theory $\Gamma\rhd\Delta$.
As a special case, we can represent the proof of a single goal $g$ by taking $\Delta=\{\vdash g\}$.
More generally, by taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem. \footnote{For example, we could represent $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
Additional data can be communicated separately; for example, the beta-reducing tactic might take a parameter specifying the maximum number of reductions to perform.

<why is it good for communicating terms and theorems? fixes ontology...>

\subsection{Isolating tactics from their home systems}
There are two approaches to obtaining a standalone tactic: either write one directly, or extract an existing tactic from a proof assistant. 
We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
<The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.>
The key requirements of the proof assistant are:
\begin{enumerate}
\item
An \OpenTheory article file reader and writer,
\item
A facility for recording proofs and converting them to use the \OpenTheory primitive inferences, and,
\item
A facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
\end{enumerate}
<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories>
<in HOL4 and HOL Light, item 2 is accomplished by augmenting the internal theorem type with a type of proofs, to track primitive inferences used. this step is made easy because of the LCF architecture>
<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>

<Implementation details to put on the web.>
<simple cgi - worth mentioning?>

@@

\section{Conclusions}
Additional (non-proof) work tactics might also do needs to be represented on another channel.

>>>>

An article file represents a theory $\Gamma\rhd\Delta$.
By taking $\Delta$ to be the theorems proved by a tactic and $\Gamma$ to be any theorems the tactic uses, we can view the result of applying a tactic as a theory.
Furthermore, we can represent a higher-order logic term in a theory by including the term in the statement of a theorem. \footnote{For example, we could represent $t$ by the theory $\{\vdash\mathtt{K\;F\;}t\}\rhd\{\vdash\mathtt{K\;F\;}t\}$, where $\mathtt{K}:\alpha\to\beta\to\alpha$ is the combinator defined by $\mathtt{K}\;x\;y=x$, and $\mathtt{F}$ is falsity (any boolean would do).
This kind of encoding is necessary to represent terms of non-boolean type, since the conclusion of a theorem must be boolean.}
Thus the same format, article files, can be used to encode terms and proofs for communication with tactics.
Additional data can be communicated separately; for example, the beta-reducing tactic might take a parameter specifying the maximum number of reductions to perform.

The main advantage to using \OpenTheory for communication is that it provides a standard, independent framework for fixing the meanings of constants between different systems. <Example?>
<Requirements~\ref{req:artr} and \ref{req:artw} above carry the additional benefit of giving a proof assistant access to traditional \OpenTheory packages (\ie not just the output of standalone tactics).>

\subsection{Isolating tactics from their home systems}
There are two approaches to obtaining a standalone tactic: either write one directly, or extract an existing tactic from a proof assistant. 
We have experimented with the second approach, taking tactics from HOL4 and from HOL Light.
The procedure is reasonably lightweight, but less flexible than writing a standalone tactic directly.
The key requirements of the proof assistant are:
\begin{enumerate}
\item
\label{req:io} an \OpenTheory article file reader and writer,
\item
\label{req:log} a facility for recording proofs and converting them to the \OpenTheory logical kernel, and,
\item
\label{req:ckpt} a facility for checkpointing the proof assistant, or for extracting a standalone executable encompassing the tactic functionality.
\end{enumerate}
<first two items are the most work, but have the additional benefit of allowing import/export of theories in opentheory repositories, as mentioned above. so: to use standalone tactics you also get to use opentheory packages, but if you can make yourself into a standalone tactic you can also export opentheory packages.>
<in HOL4 and HOL Light, item 2 is accomplished by augmenting the internal theorem type with a type of proofs, to track primitive inferences used. this step is made easy by LCF architecture: could be very difficult in systems that don't already generate proofs somehow, or with significantly different kernel (cf isabelle)>
<either use Polyml.export or checkpointing to capture a function that reads a goal article, runs the tactic (recording the proof), and writes the proof article back>

<Implementation details to put on the web: simple cgi>

@@

\section{Conclusions}
